How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of Apgar5 Scores

### 2024-11-29 paper/paper.qmd
'enhancement' parts upload

```{r}
#| label: fig-apgar1
#| fig-cap: "The distribution of APGAR5 scores across the entire observation in the original dataset **Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-apgar2
#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-apgar3
#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-numtreat
#| fig-cap: The distribution of observations across different number of treatments used during delivery and labor
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-indc1
#| fig-cap: The distribution of observations in the raw dataset based on whether the infant received induction of labor or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-indc2
#| fig-cap: The distribution of observations in the analysis dataset based on whether the infant received induction of labor or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-augmt1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-augmt2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-ster1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```
```{r}
#| label: fig-ster2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-antb1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-antb2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-chor1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```
```{r}
#| label: fig-chor2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-anes1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-anes2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false

@@

Observational data serve as the main database in biomedical research, particularly when randomized controlled trials (RCTs) are impractical due to ethical or logistical constraints [@healthknowledge2024]. These datasets, often derived from electronic health records, patient surveys, or longitudinal studies [@healthknowledge2024], provide valuable insights into real-world healthcare outcomes. However, their inherent limitations necessitate careful consideration to ensure robust and credible findings. Among the key challenges are confounding variables, selection bias, and measurement errors [@tellingstories]. For instance, in analyzing the effects of prenatal steroids on neonatal outcomes, unmeasured factors such as maternal health or socioeconomic status can introduce confounding [@natality202]. Advanced statistical techniques like propensity score matching or instrumental variables are essential to isolate causal effects from observational data [@Olier2023ObservationalData].

@@

>>>>

```{r}
#| label: fig-apgar1
#| fig-cap: "The distribution of Apgar5 scores across the entire observation in the original dataset (**Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-apgar2
#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-apgar3
#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-numtreat
#| fig-cap: "The distribution of observations across different number of treatments used during delivery and labor"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-indc1
#| fig-cap: The distribution of observations in the raw dataset based on whether the infant received induction of labor or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-indc2
#| fig-cap: The distribution of observations in the analysis dataset based on whether the infant received induction of labor or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-augmt1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-augmt2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received augmentation of labor or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-ster1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-ster2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received steroid treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-antb1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-antb2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received antibiotic treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-chor1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-chor2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received chorioamnionitis treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-anes1
#| fig-cap: The distribution observations in the raw dataset based on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

```{r}
#| label: fig-anes2
#| fig-cap: The distribution observations in the analysis datset based on whether the infant received anesthesia treatments or not
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4

@@

Observational data serve as the main database in biomedical research, particularly when randomized controlled trials (RCTs) are impractical due to ethical or logistical constraints [@healthknowledge2024]. These datasets, often derived from electronic health records, patient surveys, or longitudinal studies [@healthknowledge2024], provide valuable insights into real-world healthcare outcomes. However, their inherent limitations necessitate careful consideration to ensure robust and credible findings. Among the key challenges are confounding variables, selection bias, and measurement errors [@tellingstories]. For instance, in analyzing the effects of prenatal steroids on neonatal outcomes, unmeasured factors such as maternal health or socioeconomic status can introduce confounding [@natality2023]. Advanced statistical techniques like propensity score matching or instrumental variables are essential to isolate causal effects from observational data [@Olier2023ObservationalData].

@@

### Model API Predictions

#### API execution details

The 'Infant Health Prediction API' provides predictions for infant health status based on maternal and delivery-related factors. By accepting structured datasets in CSV format, it offers researchers a powerful tool to generate real-time predictions using advanced maching learning models.

- Endpoint URL: `http://127.0.0.1:3505/predict_csv_local`

- Description: This endpoint processes a local CSV file containing test data and provides predictions from both the Random Forest and Bayesian Linear models.

  - Random Forest Model:

    Handles complex interactions between variables to generate robust predictions.

  - Bayesian Linear Model:

    Offers interpretable results with uncertainty quantification.

#### Input requirements

- The input file must be a CSV containing only numeric columns.

- Required features include drug usage indicators and delivery-related factors.

- Missing or non-numeric data will result in validation errors.

#### Output requirements

The API returns predictions in JSON format. Each entry corresponds to a row in the input dataset, with predictions for models:

    "Random_Forest_Predictions": 5.0648,

    "Bayesian_Linear_Model_Predictions": 5.1659

The CURL command used to trigger the API:

```bash

curl -X POST 'http://127.0.0.1:3505/predict_csv_local' -H 'accept: */*' -d ''

### Shiny web applications for visualizing and analyzing predictions

The 'Infant Health Prediction App' is a Shiny web application that enables users to:
- Upload a dataset containing infant-related data in CSV format.
- Run predictions for infant health status using both a Random Forest model and a Bayesian Linear model.
- View prediction results in an interactive table.
- Explore visualizations of the predictions through histograms.

#### Features
1. Data Upload and Preview: Users can upload datasets and preview them in an interactive table.
2. Model Predictions: Predictions from both models are displayed in a table for easy comparison.
3. Visualizations: Histograms provide insights into the distribution of predictions from each model.

#### How to Run the App
To launch the app locally, use the following R command:

```r

library(shiny)

runApp("../scripts/07-shiny_app.R")

#### App Interface
Below is the interface of the Infant Health Prediction App:

![App Interface](../models/results/interface_screenshot.png)

#### Prediction Table
The predictions are displayed in an interactive table:

![Prediction Table from the Shiny Web Application](../models/results/prediction_table_screenshot.png)

#### Visualizations
The histograms show the distribution of model predictions:

![Random Forest Histogram from the Shiny Web Application](../models/results/random_forest_histogram.png)

![Bayesian Linear Model Histogram from the Shiny Web Application](../models/results/bayesian_linear_histogram.png)
